Multi-target detection with application to cryo-electron microscopy

### 2019-03-08 inverse_problems/inverse_problems.tex
sketch of summary section

 %Note that the expected total number of signals occurring in $y$ is equal to $(N - L + 1)\gamma / L$, and consequently the total lengths of all signals (including overlaps) divided by $N$ is equal to
%
%\begin{align}
%\frac{N - L + 1}{N} \cdot L \cdot \frac{\gamma  }{L} 
%\quad \overset{N\to\infty}{\longrightarrow} \quad \gamma.
%\end{align}
%
%In particular, this model allows for (additive) signal overlap

%Specifically, for some parameter $\gamma > 0$, the total number of signal occurrences is drawn from a Poisson distribution with expectation $(N-L+1)\gamma$. For each occurrence, a starting location $i$ is drawn uniformly and independently at random from $\{0, \ldots, N-L\}$. 
%Then, one of the signals $x_1, \ldots, x_K$ is selected independently at random from a fixed distribution and added to the observation $y$ with first entry positioned at $i$. 
%By definition of the Poisson process, the entries of $s$ are independent and follow a Poisson distribution with expectation $\gamma$. In particular, this model allows for (additive) signal overlap. %In Section~\TODO{ref?}, we show that the autocorrelations of $y$ are nonetheless equivalent to those arising under our well-separated model above.

 %More precisely,
 %Therefore, in the large $N$ limit $\gamma$ may be interpreted as the signal density, as in the well-separated model.

@@

Owing to the separation condition, when correlating $y$ with shifted versions of itself for shifts in $0, \ldots, L-1$, any given occurrence of $x$ in $y$ is only ever correlated with itself, and never with another occurrence. As a result, the autocorrelations of $y$ depend on the corresponding autocorrelations of $x$, the noise level $\sigma$ and the density $\gamma$ (which is a weak dependence on the support signal $s$). Specifically, we show the following identities in Section~\ref{sec:autocorrelation_computation}:
\begin{align} 
	a_y^1 & = \gamma a_{x}^1, \label{eq:mean_micrograph} \\
	a_y^2[\ell] & = \gamma a_{x}^2[\ell] + \sigma^2\delta[\ell], \label{eq:ac2_micrograph}\\
	a_y^3[\ell_1,\ell_2] & = \gamma a_{x}^3[\ell_1,\ell_2]  + \sigma^2\gamma a_{x}^1  \big(\delta[\ell_1]+\delta[\ell_2]
	+\delta[\ell_1-\ell_2]\big), \label{eq:ac3_micrograph}
\end{align}
where $\delta[0] = 1$ and $\delta[\ell \neq 0] = 0$, and indices $\ell, \ell_1, \ell_2$ are in the range $0 \leq \ell \leq L-1$. Terms proportional to $\sigma^2$ are due to noise. If $\sigma$ is known, they can be handled easily. If $\sigma$ is unknown, one can either estimate it form the data, or one can ignore the few entries of the autocorrelations that are affected by $\sigma$---one in $a_y^2$ and $3L-2$ in $a_y^3$, a relatively small number in both cases.

@@

In Section~\ref{sec:proof_prop_poisson}, we prove the following expressions for the autocorrelations of $y$ under the Poisson model. As in the well-separated model, the observed autocorrelations do not depend on individual occurrences of the signal, but only on the distribution of $x$ itself, and the parameters $\gamma$ and $\sigma$: %\TODO{Do these identities hold even without taking a limit? I guess so, but let's be sure.}
%
\begin{align}
	a_y^1  & =  \gamma a_{x}^1, \label{eq:mean_micrograph2} \\
	a_y^2[\ell] & = \gamma a_x^2[\ell] + \sigma^2 \delta[\ell] + (\gamma a_x^1)^2, \label{eq:ac2_micrograph2} \\
	a_y^3[\ell_1,\ell_2]  & =  \gamma a_x^3[\ell_1,\ell_2] + \sigma^2 \gamma a_x^1 \big(\delta[\ell_1]+\delta[\ell_2] +\delta[\ell_1-\ell_2] \big) \nonumber \\
						  & \qquad + (\gamma a_x^1)^3 + \gamma a_x^1 \cdot ( \gamma a_x^2[\ell_1] + \gamma a_x^2[\ell_2] + \gamma a_x^2[\ell_2-\ell_1]). \label{eq:ac3_micrograph2}
\end{align}
Note that from those autocorrelations, it is easy to retrieve the autocorrelations of the well-separated model for all entries unaffected by $\sigma$. If $\sigma$ is known, then it is true for all entries.

@@

%
%From these values, one may solve for the terms $\gamma a_x^1$, $\gamma a_x^2[\ell] + \sigma^2 \ell[\ell]$, and $\gamma a_x^3[\ell_1,\ell_2]$. Indeed, recovering  $\gamma a_x^1$ and $\gamma a_x^2[\ell] + \sigma^2 \ell[\ell]$ is immediate from \eqref{eq:mean_micrograph2} and \eqref{eq:ac2_micrograph2}, and $\gamma a_x^3[\ell_1,\ell_2]$ then follows from
%%
%\begin{align}
%\gamma a_x^3[\ell_1,\ell_2]
%= a_y^3[\ell_1,\ell_2] - (\gamma a_x^1)^3
%- \gamma a_x^1 \cdot \big(a_y^2[\ell_1]  
%+ a_y^2[\ell_2] + a_y^2[\ell_1-\ell_2]- 3(\gamma a_x^1)^2 \big).
%\end{align}

%\begin{corollary}[informal for now] \label{cor:ac_equivalence}
%The first and the second autocorrelations  in the well-separated and the Poisson models are equivalent. 
%If $\sigma$ is known, then it is also true for the third-order autocorrelation. If $\sigma$ is unknown, the it is true whenever $\ell_1\neq 0$, $\ell_2\neq 0$ and $\ell_1\neq \ell_2$. 
%\end{corollary}
%
%It is now easy to show how to solve for both $\gamma$ and $\sigma$ from these expressions, assuming $x$ is generic. Indeed, we observe the product
%%
%\begin{align}
%(\gamma a_x^1) (\gamma a_x^2[1])
%&=  \frac{\gamma^2}{L^2} \left( \sum_{i=0}^{L-1} x[i] \right)
%\left(\sum_{j=0}^{L-2} x[j]x[j+1] \right)
%\nonumber \\
%&= \frac{\gamma^2}{L^2} \sum_{j=0}^{L-2} \sum_{i=0}^{L-1} x[i]x[j]x[j+1]
%\nonumber \\
%&= \frac{\gamma^2}{L} \left\{ \sum_{\ell=0}^{L-1} a_x^3[1,\ell]
%+ \sum_{\ell=1}^{L-2} a_x^3[\ell,\ell+1]  \right\}.
%\end{align}
%%
%Since we also observe $\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] + \sum_{\ell=1}^{L-2} \gamma a_x^3[\ell,\ell+1]$, \TODO{Is there a bias?} we can form the ratio and solve for $\gamma$: 
%
%\begin{align}
%\gamma = L \frac{(\gamma a_x^1) (\gamma a_x^2[1])}
%{\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] 
%	+ \sum_{\ell=1}^{L-2} \gamma a_x^3[\ell,\ell+1]}.
%\end{align}
%%With $\gamma$ in hand, we can estimate $\sigma^2$:
%%
%\begin{align}
%\sigma^2 = a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\frac{L (a^1_y)^2}{\gamma}
%- (2 L - 1) (a_y^1)^2.
%\end{align}
%
\begin{comment}
In this section we omit the affect of the noise as it remains the same for the well-separated model. We aim to show that the moments of

Let us denote by $m_Y^q$ the moments of $Y$:
%
\begin{align}
%
m_X^1[i] &= \E X[i], \quad 0 \le i \le L-1,   \nonumber\\ 
m_X^2[i,j] &= \E X[i] X[j], \quad 0 \le i,j \le L-1, \\ 
m_X^3[i,j,k] &= \E X[i] X[j] X[k], \quad 0 \le i,j,k \le L-1. \nonumber
\end{align}
%
Note that 
\begin{align}
a_X^1 &= \frac{1}{L}\sum_{i=0}^{L-1}m_X^1[i], \nonumber \\
a_X^2[\ell] &= \frac{1}{L}\sum_{i=0}^{L-1}m_X^2[i,i+\ell], \\
a_X^3[\ell_1,\ell_2] &= \frac{1}{L}\sum_{i=0}^{L-1}m_X^3[i,i+\ell_1,i+\ell_2]. \nonumber
\end{align}

In the following,  we show that the moments of the observed data under the Poisson process can be written in terms of the autocorrelations of the well-separated model.

\begin{proposition} \label{prop:poisson}
	For any $i$,  the moments under the Poisson process are equal to 
	\begin{align}
	m_y^1[i] &= \gamma L a_X^1 \nonumber \\
	m_y^2[i,i+\ell] &= (\gamma a_X^1)^2 + \gamma a_X^2[\ell],\quad \ell=0,\ldots,L-1, \\
	m_y^3[i,i+\ell_1,i+\ell_2] &= (\gamma a_X^1)^3 + \gamma a_X^1  \cdot ( \gamma a_X^2[\ell_1] + \gamma a_X^2[\ell_2] + \gamma a_X^2[\ell_2-\ell_1]) + \gamma a_X^3[\ell_1,\ell_2]. \nonumber  
	\end{align}
\end{proposition}

\begin{corollary} \TODO{re-write}
	From the first three moments of the Poisson process model, one can recover the first three moments from the strongly-separated model.
\end{corollary}
\end{comment}
%
%\TODO{What we want to have in this section}
%\begin{enumerate}
%	\item Recall the Poisson model (it's already described at the end of intro, so keep it short).
%	\item Define a proper notion of $\gamma$ (careful with scaling $L$).
%	\item State formulas for $a_y^q$ with $q = 1, 2, 3$ (ideally, including noise terms) to parallel what we did in the well-separated section. It would be better to do everything with autocorrelations rather than moments to keep the story uniform.
%	\item (optional) Claim: under some assumptions (knowing $\gamma$? Others?), the autocorrelations of $y$ of order $q = 1, 2, 3$ under the Poisson model provide the same information about $x$ as the corresponding autocorrelations under the well-separated model. Could also be in Section 4.1.
%\end{enumerate}
%

@@

First, we observe that if  the noise level $\sigma$ is known, generally, one can estimate $\gamma$ from the first two moments of the micrograph.
The other direction is true as well: if $\gamma>0$ is known, then one can estimate $\sigma^2$.
 The proof is provided in Section~\ref{sec:proof_prop_gamma}.

@@

Using third-order autocorrelation information of $y$, both the ratio $\gamma$ and the noise $\sigma$ can be determined simultaneously. For the following results, when we say that a result holds for a ``generic'' signal $x$, we mean that the set of signals which cannot be determined by these measurements
has Lebesgue measure zero. 
In particular, this means that we can recover
almost all signals with the given measurements. The proof is provided in Section~\ref{sec:proof_prop_gamma_sigma}.

@@

Similarly to the homogeneous well-separated model, the observed autocorrelations under the Poisson model identify $x,\gamma$ and $\sigma^2$ uniquely.
Opposed to Proposition~\ref{prop:gamma_sigma}, these quantities can be computed explicitly. 
In Section~\ref{sec:proof_uniqueness_poisson} we prove the following result:

@@

In this section, we present three numerical experiments. In the first two experiments, we consider the heterogeneous model~\eqref{eq:model}. First, with a fixed noise level, we show how the estimation quality improves as the length of the observation grows. Second, we explore how many signals can be recovered as a function of their length, in an infinite data regime where effects of the noise have been averaged out. In the last experiment, we extend our model to 2-D signals.
The code for all experiments is available at \url{https://github.com/PrincetonUniversity/BreakingDetectionLimit}.

@@

Autocorrelation analysis can be carried out in dimensions greater than one. In the following experiment, we estimate 
a 50-by-50 pixel grayscale picture of Einstein with mean zero from a growing number of observations $y$. Each observation is of size $4096\times 4096$ pixels and contains 700 occurrences on average at random locations, while maintaining the separation condition~\eqref{eq:spacing} in each axis separately.
The observations are contaminated with additive white Gaussian noise with standard deviation $\sigma = 3$, illustrated in Figure~\ref{fig:micro_example}. 
 % corresponding  to SNR = ${\frac{M\|x\|_F^2} {\sigma^2N} \approx1/370}$. 
%As explained in Section~\ref{sec:high_dimensions}, an image is determined uniquely form its second-order autocorrelation.
%Hence, to use this property and keep things simple, we assume the number of signal occurrences and the noise standard deviation are known so that we can estimate the second-order autocorrelation of the image from thos the observation directly.  %Micrographs are generated such that any two occurrences are always separated by at least 49 pixels in each direction in accordance with the separation condition~\eqref{eq:spacing}.

@@

\begin{figure}[t]
	\centering
	\begin{subfigure}[h]{0.33\linewidth}
		\centering
		\includegraphics[width=.8\linewidth]{micrograph_Einstein_example_clean}
		\caption{$\sigma = 0$}
	\end{subfigure}%
	\begin{subfigure}[h]{0.33\linewidth}
		\centering
		\includegraphics[width=.8\linewidth]{micrograph_Einstein_example_s05}
		\caption{$\sigma = 0.5$}
	\end{subfigure}
	\begin{subfigure}[h]{0.33\linewidth}
		\centering
		\includegraphics[width=.8\linewidth]{micrograph_Einstein_example_s3}
		\caption{$\sigma = 3$}
	\end{subfigure}
	\caption{\label{fig:micro_example} Example of observations for the 2-D experiment (of size $250\times 250$) with additive white Gaussian noise of variance $\sigma^2$ for increasing values of $\sigma$. Each observation contains the same four occurrences of a $50 \times 50$ image of Einstein. In panel (c), the noise level is such that it is challenging to detect  the planted images.} %In fact, it can be shown that at low SNR, reliable detection of individual image occurrences is impossible, even if the true image is known. } 	
\end{figure}
%By analogy to cryo-EM, this depicts a scenario where particle picking cannot be done.}
%\vspace{-8pt}

@@

\section{Proofs}

%Proofs for some of the results presented in Section~\ref{sec:AC_analysis} and~\ref{sec:extensions} follow in this section.

\subsection{Autocorrelations in the well-separated model} \label{sec:autocorrelation_computation}

%Let us define
%\begin{equation*}
%\gamma = \frac{|s|L}{N}>0.
%\end{equation*}
%Then the first order autocorrelation of the observation is given by 
%\begin{equation*}
%a_y^1 = \E_y\left\{\frac{1}{N}\sum_{i=0}^{N-1} y[i]\right\} =
%\frac{|s|}{N}\E_x\left\{\sum_{i=0}^{L-1}x[i]\right\} +    
%\E_{\varepsilon}\left\{\frac{1}{N}\sum_{i=0}^{N-1}\varepsilon[i]\right\}
%=\gamma a_x^1,
%\end{equation*}
%where $\E_z$ means expectation with respect to the distribution of $z$.
%%the noise term converges to zero almost surely (a.s.)\ by the strong law of large numbers.
%
%We proceed with the second-order autocorrelation for  $\ell\in\{0,\ldots,L-1\}$. We can compute:
%%
%\begin{align*}
%%
%a_y^2[\ell] & = \E_y\left\{\frac{1}{N}\sum_{i=0}^{N-1-\ell}y[i]y[i+\ell]\right\}
%\nonumber \\
%& = \frac{|s|}{N}\E_x\left\{\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]\right\}  + \E_{\varepsilon}\left\{\frac{1}{N}\sum_{i=0}^{N-1-\ell}\varepsilon[i]\varepsilon[i+\ell] \right\},
%%
%\end{align*}
%where all cross terms vanish in expectation owing to independence of $x$ of $\varepsilon$, and $\E{\varepsilon}=0$.  
%Hence, 
%\begin{align} \label{eq:2nd_moment_signal_term}
%%
%a_y^2[\ell] & = \gamma a_x^2[\ell] + \frac{N-1-\ell}{N}\sigma^2\ell[\ell],
%\end{align}
%where $\ell[0]=1$ and zero otherwise. 
%
%We now analyze the third-order autocorrelation. 
%Let $x_{(j)}$ denote the $j$th
%realization of $x$, with starting position $s_{(j)}$, for $j=1,\ldots,|s|$.
%Let us fix $0\leq \ell_2\leq\ell_1\leq L-1$. We have:
%\begin{align*}
%%
%a_y^3[\ell_1,\ell_2] 
%=& \E_y\left\{\frac{1}{N}\sum_{i=0}^{N-1-\ell_1} y[i]y[i+\ell_1]y[i+\ell_2]\right\}
%\\
%%
%\stackrel{(1)}=& \E_{x_{(1)},\ldots,x_{(|s|)}}\left\{ \frac{|s|L}{N}\frac{1}{|s|}\sum_{j=1}^{|s|} 
%	\frac{1}{L}\sum_{i=0}^{L-1-\ell_1}x_{(j)}[i]x_{(j)}[i+\ell_1]x_{(j)}[i+\ell_2]  \right\} \\
%\stackrel{(2)}+& \E_\varepsilon\left\{\frac{1}{N}\sum_{i=0}^{N-1-\ell_1} \ep[i]\ep[i+\ell_1]\ep[i+\ell_2]\right\}
%\ \\
%\stackrel{(3)}+& \E_{x_{(1)},\ldots,x_{(|s|)},\varepsilon}\left\{\frac{1}{N}\sum_{j=1}^{|s|} 
%	\sum_{i=0}^{L-1} x_{(j)}[i]\ep[s_{(j)} + i+\ell_1]\ep[s_{(j)}+ i+\ell_2]\right\} \\
%\stackrel{(4)}+& \E_{x_{(1)},\ldots,x_{(|s|)},\varepsilon}\left\{\frac{1}{N}\sum_{j=1}^{|s|} 
%	\sum_{i=0}^{L-1} \ep[s_{(j)}+i-\ell_1]x_{(j)}[i]\ep[s_{(j)}+ i+\ell_2-\ell_1]\right\}
%\\
%\stackrel{(5)}+& \E_{x_{(1)},\ldots,x_{(|s|)},\varepsilon}\left\{\frac{1}{N}\sum_{j=1}^{|s|} 
%	\sum_{i=0}^{L-1} \ep[s_{(j)}+i-\ell_2]\ep[s_{(j)}+i+\ell_1-\ell_2]x_{(j)}[i]\right\}
%\\ \stackrel{(6)}+ & \E_{x_{(1)},\ldots,x_{(|s|)},\varepsilon}\left\{\frac{1}{N}\sum_{j=1}^{|s|} 
%	\sum_{i=0}^{L-\ell_1+\ell_2-1} \ep[s_{(j)}+i-\ell_2]x_{(j)}[i+\ell_1-\ell_2]x_{(j)}[i]\right\}
%\\
%\stackrel{(7)}+ & \E_{x_{(1)},\ldots,x_{(|s|)},\varepsilon}\left\{\frac{1}{N}\sum_{j=1}^{|s|} 
%	\sum_{i=0}^{L-\ell_2-1} x_{(j)}[i]\ep[s_{(j)} + i+\ell_1]x_{(j)}[i+\ell_2]\right\}
%\\ \stackrel{(8)}+ & \E_{x_{(1)},\ldots,x_{(|s|)},\varepsilon}\left\{\frac{1}{N}\sum_{j=1}^{|s|} 
%	\sum_{i=0}^{L-\ell_1-1} x_{(j)}[i]x_{(j)}[i+\ell_1]\ep[s_{(j)}+ i+\ell_2]\right\}.
%%
%\end{align*}
%Terms (6), (7) and (8) are linear in $\ep$ and  thus zero.
%Term (1) is equal  to $\gamma a_x^3[\ell_1,\ell_2]$  for the same reasons as~\eqref{eq:2nd_moment_signal_term}. 
%Term (2) is always zero by properties of Gaussian variables. 
%To deal with terms (3)--(5), we must distinguish between different values of $\ell_1$ and $\ell_2$.
%
%Term (3) is always zero, unless $\ell_1=\ell_2$, so that it is equal to 
%$\gamma\sigma^2 a_x^1 \ell[\ell_1-\ell_2]$.
%The same goes for Terms (4) and (5), which are equal, respectively, to
%$\gamma\sigma^2 a_x^1 \ell[\ell_2]$ 
%and $\gamma\sigma^2 a_x^1 \ell[\ell_1]$.

@@

\subsection{Autocorrelations in the Poisson model} \label{sec:proof_prop_poisson}

%Throughout this section, we will use the capital letter $X$ to represent the random unobserved vector of length $L$, the capital letter $Y$ to represent the random observed vector of length $N$, and the capital letter $S$ to represent the unobserved vector of occurrences.

@@

%

\subsubsection{Computing $a_y^1$}

@@

%

\subsubsection{Computing $a_y^2$}

@@

%

\subsubsection{Computing $a_x^3$}

@@

%
%\TODO{Needs a complete rewrite to match the paper, in particular the modified notation and the fact we only work with autocorrelations now (not models).}
%
%\subsubsection{First moment}
%
%To compute the first moment of $y$, we will first condition on $M = (M_1,\dots,M_{N-L+1})$, and then average over $M$. We have:
%%
%\begin{align}
%%
%\E[y[i] | M] = \sum_{j=0}^{L-1} \sum_{k=1}^{M_{i-j}} \E X_k^{i-j}[j]
%= \sum_{j=0}^{L-1} \sum_{k=1}^{M_{i-j}} m_y^1[j]
%= \sum_{j=0}^{L-1} M_{i-j} m_y^1[j].
%%
%\end{align}
%%
%Now taking expectations over $M$ we see:
%%
%\begin{align}
%%
%\E Y[i] = \gamma \sum_{j=0}^{L-1}  m_y^1[j] = \gamma La_x^1.
%%
%\end{align}
%
%
%%
%
%
%\subsubsection{Second moment}
%
%Again, we will condition on $M$ first, and then take the expectation over $M$. Fix $i_1 \ne i_2$, and let $\ell = i_2 - i_1$. Then:
%%
%\begin{align}
%%
%Y_{i_1} Y_{i_2} 
%&= \sum_{j_1=0}^{L-1} \sum_{j_2=0}^{L-1} 
%\sum_{k_1=1}^{M_{i_1-j_1}}\sum_{k_2=1}^{M_{i_2-j_2}}
%X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_2 - j_2}[j_2].
%%
%\end{align}
%
%We break up the double sum over $j_1$ and $j_2$ into two terms: one where $j_2 \ne j_1 + \ell$, and one where $j_2 = j_1 + \ell$ or equivalently $i_1-j_1 = i_2-j_2$. In the first case, all the terms are independent, and so the expectation factors. In the second case, when $k_1 \ne k_2$ we have independence, but otherwise not. This gives (all expectations are conditional on $M$):
%%
%\begin{align} \label{moment2-condm}
%%
%\E Y_{i_1} Y_{i_2}
%=& \sum_{j_1=0}^{L-1} \sum_{j_2=0}^{L-1} 
%\sum_{k_1=1}^{M_{i_1-j_1}}\sum_{k_2=1}^{M_{i_2-j_2}}
%\E X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_2 - j_2}[j_2]
%\nonumber \\
%=& \sum_{j_1 - j_2 \ne \ell} \sum_{k_1} 
%\sum_{k_2} \E X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_2 - j_2}[j_2]
%\nonumber \\
%& + \sum_{j_1 = 0}^{L-1} \sum_{k_1 \ne k_2} 
%\E X_{k_1}^{i_1-j_1}[j_1] X_{k_2}^{i_1 - j_1}[j_1+\ell]
%\nonumber \\
%& + \sum_{j_1 = 0}^{L-1} \sum_{k_1=1}^{M_{i_1-j_1}} 
%\E X_{k_1}^{i_1-j_1}[j_1] X_{k_1}^{i_1-j_1}[j_1 + \ell] 
%\nonumber \\
%=& \sum_{j_1 - j_2 \ne \ell} M_{i_1-j_1} M_{i_2 - j_2} \M_1[j_1] \M_1[j_2]
%\nonumber \\
%& + \sum_{j_1 = 0}^{L-1} M_{i_1-j_1}(M_{i_1-j_1} - 1) \M_1[j_1] \M_1[j_1 + \ell]
%\nonumber \\
%& + \sum_{j_1 = 0}^{L-1} M_{i_1-j_1} \M_2[j_1,j_1 + \ell] .
%%
%\end{align}
%%
%Now take expectations over the Poisson random variables, using this fact:
%%
%\begin{lem} \label{lem-choose}
%	If $M \sim \Poisson(\gamma)$, then 
%	\begin{align}
%	%
%	\E {M\choose k} = \frac{\gamma^k}{k!}.
%	%
%	\end{align}
%\end{lem}
%
%
%We get (now the expectation is over $M$ and $X$):
%%
%\begin{align}
%%
%\E Y_{i_1} Y_{i_2} 
%=& \sum_{j_1 - j_2 \ne \ell} \E M_{i_1-j_1} M_{i_2 - j_2} \M_1[j_1] \M_1[j_2]
%\nonumber \\
%& + \sum_{j_1 = 0}^{L-1} \E M_{i_1-j_1}(M_{i_1-j_1} - 1) \M_1[j_1] \M_1[j_1 + \ell]
%\nonumber \\
%& + \sum_{j_1 = 0}^{L-1} \E M_{i_1-j_1} \M_2[j_1,j_1 + \ell]
%\nonumber \\
%=& \sum_{j_1 - j_2 \ne \ell} \gamma^2 \M_1[j_1] \M_1[j_2]
%+ \sum_{j_1 = 0}^{L-1} \gamma^2 \M_1[j_1] \M_1[j_1 + \ell]
%\nonumber \\
%& + \sum_{j_1 = 0}^{L-1} \gamma \M_2[j_1,j_1 + \ell]
%\nonumber \\
%=&  \bigg(\gamma \sum_{j = 0}^{L-1} \M_1[j] \bigg)^2
%+ \gamma \sum_{j = 0}^{L-1} \M_2[j,j + \ell]
%\nonumber \\
%=&  (\gamma \L_1)^2 + \gamma \L_2(\ell).
%%
%\end{align}
%
%But the first term in the sum is just the square of the first moment of $Y$; so from the first two moments we can recover $\gamma \L_2(\ell)$, which is just the expected power spectrum of the random vector $X$, i.e.\ the usual second moment we have been working with.
%
%
%%
%
%
%\subsubsection{Third moment}
%

>>>>

@@

Owing to the separation condition, when correlating $y$ with shifted versions of itself for shifts in $0, \ldots, L-1$, any given occurrence of $x$ in $y$ is only ever correlated with itself, and never with another occurrence. As a result, the autocorrelations of $y$ depend on the corresponding autocorrelations of $x$, the noise level $\sigma$ and the density $\gamma$ (which is a weak dependence on the support signal $s$). Specifically, we show the following identities in Appendix~\ref{sec:autocorrelation_computation}:
\begin{align} 
	a_y^1 & = \gamma a_{x}^1, \label{eq:mean_micrograph} \\
	a_y^2[\ell] & = \gamma a_{x}^2[\ell] + \sigma^2\delta[\ell], \label{eq:ac2_micrograph}\\
	a_y^3[\ell_1,\ell_2] & = \gamma a_{x}^3[\ell_1,\ell_2]  + \sigma^2\gamma a_{x}^1  \big(\delta[\ell_1]+\delta[\ell_2]
	+\delta[\ell_1-\ell_2]\big), \label{eq:ac3_micrograph}
\end{align}
where $\delta[0] = 1$ and $\delta[\ell \neq 0] = 0$, and indices $\ell, \ell_1, \ell_2$ are in the range $0 \leq \ell \leq L-1$. Terms proportional to $\sigma^2$ are due to noise. If $\sigma$ is known, they can be handled easily. If $\sigma$ is unknown, one can either estimate it form the data, or one can ignore the few entries of the autocorrelations that are affected by $\sigma$---one in $a_y^2$ and $3L-2$ in $a_y^3$, a relatively small number in both cases.

@@

In Appendix~\ref{sec:proof_prop_poisson}, we prove the following expressions for the autocorrelations of $y$ under the Poisson model. As in the well-separated model, the observed autocorrelations do not depend on individual occurrences of the signal, but only on the distribution of $x$ itself, and the parameters $\gamma$ and $\sigma$: 
%
\begin{align}
	a_y^1  & =  \gamma a_{x}^1, \label{eq:mean_micrograph2} \\
	a_y^2[\ell] & = \gamma a_x^2[\ell] + \sigma^2 \delta[\ell] + (\gamma a_x^1)^2, \label{eq:ac2_micrograph2} \\
	a_y^3[\ell_1,\ell_2]  & =  \gamma a_x^3[\ell_1,\ell_2] + \sigma^2 \gamma a_x^1 \big(\delta[\ell_1]+\delta[\ell_2] +\delta[\ell_1-\ell_2] \big) \nonumber \\
						  & \qquad + (\gamma a_x^1)^3 + \gamma a_x^1 \cdot ( \gamma a_x^2[\ell_1] + \gamma a_x^2[\ell_2] + \gamma a_x^2[\ell_2-\ell_1]). \label{eq:ac3_micrograph2}
\end{align}
Note that from those autocorrelations, it is easy to retrieve the autocorrelations of the well-separated model for all entries unaffected by $\sigma$. If $\sigma$ is known, then it is true for all entries.

@@

@@

First, we observe that if  the noise level $\sigma$ is known, generally, one can estimate $\gamma$ from the first two moments of the micrograph.
The other direction is true as well: if $\gamma>0$ is known, then one can estimate $\sigma^2$.
 The proof is provided in Appendix~\ref{sec:proof_prop_gamma}.

@@

Using third-order autocorrelation information of $y$, both the ratio $\gamma$ and the noise $\sigma$ can be determined simultaneously. For the following results, when we say that a result holds for a ``generic'' signal $x$, we mean that the set of signals which cannot be determined by these measurements
has Lebesgue measure zero. 
In particular, this means that we can recover
almost all signals with the given measurements. The proof is provided in Appendix~\ref{sec:proof_prop_gamma_sigma}.

@@

Similarly to the homogeneous well-separated model, the observed autocorrelations under the Poisson model identify $x,\gamma$ and $\sigma^2$ uniquely.
Opposed to Proposition~\ref{prop:gamma_sigma}, these quantities can be computed explicitly. 
In Appendix~\ref{sec:proof_uniqueness_poisson} we prove the following result:

@@

In this section, we present three numerical experiments. In the first two experiments, we consider the heterogeneous model~\eqref{eq:model}. First, with a fixed noise level, we show how the estimation quality improves as the length of the observation grows. Second, we explore how many signals can be recovered as a function of their length, in an infinite data regime where effects of the noise have been averaged out. In the last experiment, we extend our model to 2-D signals.
We run the experiments on a shared computer with 144 logical CPUs of type Intel(R) Xeon(R)
CPU E7-8880 v3 $@$ 2.30GHz and 755Gb of RAM; we use at most 72 of these CPUs.
The code for all experiments is available at \url{https://github.com/PrincetonUniversity/BreakingDetectionLimit}.

@@

Autocorrelation analysis can be carried out in dimensions greater than one. In the following experiment, we estimate 
a 50-by-50 pixel grayscale picture of Einstein with mean zero from a growing number of observations $y$. Each observation is of size $4096\times 4096$ pixels and contains 700 occurrences on average at random locations, while maintaining the separation condition~\eqref{eq:spacing} in each axis separately.
The observations are contaminated with additive white Gaussian noise with standard deviation $\sigma = 3$, illustrated in Figure~\ref{fig:micro_example}.

@@

\begin{figure}[t]
	\centering
	\begin{subfigure}[h]{0.33\linewidth}
		\centering
		\includegraphics[width=.8\linewidth]{micrograph_Einstein_example_clean}
		\caption{$\sigma = 0$}
	\end{subfigure}%
	\begin{subfigure}[h]{0.33\linewidth}
		\centering
		\includegraphics[width=.8\linewidth]{micrograph_Einstein_example_s05}
		\caption{$\sigma = 0.5$}
	\end{subfigure}
	\begin{subfigure}[h]{0.33\linewidth}
		\centering
		\includegraphics[width=.8\linewidth]{micrograph_Einstein_example_s3}
		\caption{$\sigma = 3$}
	\end{subfigure}
	\caption{\label{fig:micro_example} Example of observations for the 2-D experiment (of size $250\times 250$) with additive white Gaussian noise of variance $\sigma^2$ for increasing values of $\sigma$. Each observation contains the same four occurrences of a $50 \times 50$ image of Einstein. In panel (c), the noise level is such that it is challenging to detect  the planted images.} 
\end{figure}

@@

\section{Summary}

This paper suggests a computational framework for estimating a signal (signals) from their multiple occurrences in noisy data based on autocorrelation analysis. 
The autocorrelations can be estimated accurately in any new level, given sufficiently long measurement. 
Given accurate estimates of the autocorrelations, the problem reduces to estimate a signal from their autocorrelation, that is, solving a system of low-order polynomials equations. The crux of the method lies in the distinctions between parameters that we aim to estimate (the signal) and the nuisance variables (individual locations of signals' occurrences). In particular, we show that the signals can be estimated, although we cannot detect the occurrences in the data.

The prime motivation for this project arises from the problem of reconstructing 3-D of macromolecules using cryo-EM.
Currently, small molecules cannot be elucidated using cryo-EM; the signal-to-noise ratio on the micrograph is too low and thus it is impossible to detection the tomographic projections. However, the goal of the cryo-EM is not to detect these projections but rather to estimate the 3-D itself; the locations of the projections, as well as the viewing direction associated with each projection, are nuisance variables. Recently, we proposed to estimate the 3-D structure directly from the micrograph, without particle picking~\cite{bendory2018toward}. 
Currently, the reconstructed volumes are of low-resolutions. It is not clear whether it is possible to achieve higher resolution without considering higher autocorrelation, which in turn will increase the sample complexity of the problem. 
Application on cryo-EM data will also require incorporating additional parameters into the model: the distribution of 3-D orientation of the molecules in the ice, the noise power spectrum (usually colored noise), the contrast-transfer function, and the distribution of the projection across the   micrograph.
These parameters will be encoded within the autocorrelations.
Then, the goal would be to estimate these parameters, along with the 3-D structure, by solving the polynomial system of equations.

\appendix

\section{Autocorrelations in the well-separated model} \label{sec:autocorrelation_computation}

@@

\section{Autocorrelations in the Poisson model} \label{sec:proof_prop_poisson}

@@

\subsection{Computing $a_y^1$}

@@

\subsection{Computing $a_y^2$}

@@

\subsection{Computing $a_x^3$}

@@

\section{Proof of Proposition~\ref{prop:gamma}} \label{sec:proof_prop_gamma}
Refer to equations \eqref{eq:mean_micrograph}--\eqref{eq:ac3_micrograph} for expressions relating the moments of $y$ and those of $x$, and the parameters $\gamma$ and $\sigma$. First,  note that 
\begin{align*}
(a^1_y)^2 & = \frac{\gamma^2}{L^2}\sum_{i=0}^{L-1}\sum_{j=0}^{L-1}x[i]x[j].
\end{align*}
Similarly, using $a_x^2[-\ell] = a_x^2[\ell]$:
\begin{align*}
2\sum_{\ell = 1}^{L-1}a_y^2[\ell] & = \gamma \sum_{\ell = 1}^{L-1} a_x^2[\ell] + a_x^2[-\ell] \\
& = \frac{\gamma}{L}\sum_{i=0}^{L-1}\sum_{\ell = 1}^{L-1} x[i]\left(x[i+\ell] + x[i-\ell]\right) \\ &= \frac{\gamma}{L} \sum_{i = 0}^{L-1} \sum_{j = 0, j \neq i}^{L-1} x[i] x[j],
\end{align*}
where the last equality is obtained by noting that, given the summation bounds, the set of pairs $(i, i\pm\ell)$ and $(i, j)$ are the same over the valid range $\{0, \ldots, L-1\}^2$. To conclude, notice that $a_y^2[0]=\frac{\gamma}{L}\sum_{i=0}^{L-1}x[i]^2 + \sigma^2$ and combine.

\section{Proof of Proposition~\ref{prop:gamma_sigma}} \label{sec:proof_prop_gamma_sigma}
